---
title: "Lab 1 - Quarto Warmup & STAT 331 Review"
author: "Sydney Potkey"
format: html
embed-resources: true
editor: source
execute: 
  echo: true
  warnings: false
---

```{r}
#| label: packages-setup

library(tidyverse)
library(ggridges)
```

## Quarto

First, let's make sure you know how to use Markdown formatting to style a Quarto
document.

1.  **Make this text bold.**

2.  *Make this text italicized.*

3.  Make these into a bullet point list:

 *Apples 
 *Bananas 
 *Potatoes

4.  Edit the YAML to remove warning messages from being output in the rendered
HTML file

5.  Using code chunk options, make it so this chunk shows the plot but not the
source code:

```{r}
#| echo: false

ggplot(data = mpg, 
       mapping = aes(y = manufacturer, x = hwy)) + 
  geom_boxplot() +
  labs(x = "",
       y = "", 
       title = "Highway Milage (mpg) for Different Car Manufacturers"
       )
```

6.  Using code chunk options, remove the messages about bandwidth
`geom_density_ridges()` chose to use:

```{r}
#| message: false
ggplot(data = mpg, 
       mapping = aes(y = manufacturer, x = hwy)) + 
  geom_density_ridges() +
  labs(x = "",
       y = "", 
       title = "Highway Milage (mpg) for Different Car Manufacturers"
       )
```

7.  Using code chunk options, make it so that these plots are printed
side-by-side:

```{r}
#| fig-ncol: 2


ggplot(data = mpg, 
       mapping = aes(y = manufacturer, x = hwy)) + 
  geom_boxplot() +
  labs(x = "",
       y = "", 
       title = "Highway Milage (mpg) for Different Car Manufacturers"
       )

ggplot(data = mpg, 
       mapping = aes(y = manufacturer, x = hwy)) + 
  geom_density_ridges() +
  labs(x = "",
       y = "", 
       title = "Highway Milage (mpg) for Different Car Manufacturers"
       )

```

8.  Using code chunk options, make it so this chunk shows the code but not the
output:

```{r}
#| eval: false

2 + 2
```

9.  Using code chunk options, make it so the file can still knit even though
this chunk has an error

```{r}
#| error: true

2 + a
```

10. Using code chunk options, create a descriptive `label` for each of the code
chunks above.

## Data Wrangling Review

Since you've already seen some `ggplot`s, let's do a bit of review on data 
handling. In this class, we will exclusively make use of tools from the 
`tidyverse` suite of packages to perform our data cleaning and wrangling
operations. If you are less familiar with these packages or it's been some time
since you used them, I would strongly recommend referencing the function 
documentation!  

For these problems, we will continue to work with the `mpg` data frame, making
various changes to the data to clean it up. 

11. The `fl` variable describes the type of fuel for each car, with levels: `p`, 
`r`, `e`, `d`, and `c`. Do some research into what each of these labels mean! 
Then, use the `if_else()` function to create a new variable (`fuel_type`) with
two levels: `petrol` (any car using petrolium-based gas) and 
`alternative energy` (any car not using petrolium-based gas). 

```{r}

glimpse(mpg)
mpg$fl

mpg |>
  mutate(fuel_type = if_else(
    fl == "p",
    "petrol",
    "alternative energy"))





```

12. The `drv` variable describes if the car has front drive (`f`), rear drive
(`r`), or four wheel drive (`4`). Let's make better labels for these values! 
Specifically, use the `case_when()` function to change the `drv` varible to have
the following levels: `front`, `rear`, `four wheel`. 

```{r}
mpg |>
  mutate(drv = case_when(
    drv == "f" ~ "front",
    drv == "r" ~ "rear",
    drv == "4" ~ "four wheel"
  ))


```

(string split + remove extra ")" )
13. The `trans` variable contains two pieces of information, (1) the
transmission style (`auto` or `manual`) and the specific type of transmission
(e.g., `l5`, `m5`). Using the `str_split()` function, create a new variable
(`trans_type`) containing the specific type of transmission of each car. Once
you've made this new variable, use the `rename()` function to change the name
of the `trans` column to `trans_style`.   
*Hint: You will need to deal with the stray parenthesis!* 

```{r}

mpg |> 
  mutate(
    trans_split = str_split(trans, "\\(", n = 2),  
    trans = map_chr(trans_split, 1),  
    trans_type = map_chr(trans_split, ~str_remove(.[2], "\\)")) 
  ) |>
  select(-trans_split) |>
  rename(trans_style = trans)



```

## Getting to know your classmates

14. Find someone who took Stat 331 from a different professor than you. Compare
your experiences. Tell me their name and professor. List one or two things that
you think you learned more about, and one or two things that they learned more
about.

I talked to Lucas who took 331 with Dr.Lund. A major difference between our classes was that his class did not use markdown or quarto for their assignments. His class also focused more on regression and statistical analysis n r than data wrangling, visualization, and function writing. 


15. Find someone in the class who **does not** share your birth month. Tell me
their name and birthday, and use R to find out how many days apart your
birthdays are.

```{r}
library(lubridate)
start_date <- ymd("2024-08-30")
end_date <- ymd("2024-09-22")

end_date - start_date




```

Zach's birthday is August 30th. Mine is September 22nd. There are 23 days between our birthdays.




